---
title: "Assignment 7"
author: "Douglas Barley"
date: "10/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(XML)
library(RCurl)
library(rlist)
library(jsonlite)
```

## Creation of three file types: HTML, XML, JSON

Using the five required textbooks for this course, there are three files respectively in HMTL, XML and JSON format in my GitHub repo <a href="https://github.com/douglasbarley/DATA607>here<a>. The files are respectively named "FiveBooksHTML.html", "FiveBooksXML.xml", and "FiveBooksJSON.json".


## Reading the files into R dataframes

Each file is read into a dataframe.

I could not figure out how to read an HTML file into R.

```{r }

HTML_URL <- "https://raw.githubusercontent.com/douglasbarley/DATA607/master/FiveBooksHTML.html"

# What to do???
  
```

Nor could I figure out how to read an XML file into R.

```{r }
XML_URL <- "https://raw.githubusercontent.com/douglasbarley/DATA607/master/FiveBooksXML.xml"

# What to do???

```

I could figure out how to read a JSON file into R.

```{r }
JSON_URL <- "https://raw.githubusercontent.com/douglasbarley/DATA607/master/FiveBooksJSON.json"

df_JSON <- fromJSON(JSON_URL)
df_JSON
```

