---
title: "Week 3 Assignment"
author: "Douglas Barley"
date: "9/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
```

### Question 1
Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"
```{r}
dataURL <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
majors <- read.csv(dataURL)
grep(pattern = 'DATA',majors$Major, value = TRUE, ignore.case = TRUE)
grep(pattern = 'STATISTICS',majors$Major, value = TRUE, ignore.case = TRUE)
```

### Question 2
Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
fruits <- data.frame(c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry"))
names(fruits) <- "fruit"

cat(paste(fruits, collapse = ","))

```

The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:

#3 Describe, in words, what these expressions will match:

`(.)\1\1`

`"(.)(.)\\2\\1"`

`(..)\1`

`"(.).\\1.\\1"`

`"(.)(.)(.).*\\3\\2\\1"`

```{r}
test1 <- list('"amana"', '"redder"', '"a\1\1"', '"anna"', 'sasquatch\1')
str_view(test1,'(.)\1\1', match = TRUE)
str_view(test1,'"(.)(.)\\2\\1"', match = TRUE)
str_view(test1,'(..)\1', match = TRUE)
str_view(test1,'"(.).\\1.\\1"', match = TRUE)
str_view(test1,'"(.)(.)(.).*\\3\\2\\1"', match = TRUE)
```

In order,
expression 1, `(.)\1\1`, accepts a first character group `(.)` followed by two consecutive backreferences to the character group `\1\1`. The first backreference will search through the string in the character group one character at a time until a character is repeated, then the second backreference will search through the string again until it finds the first backreference, which matches it. This will match a string like `a\1\1` in that the character group will contain the letter "a" which will be found by the first backreference as a match, and the second backreference will also find the character "a" in the character group, thus terminating the regex.

expression 2, `"(.)(.)\\2\\1"`, accepts a first character group `(.)` followed by a second character group `(.)` followed by a backreference to the second character group `\\2` finishing with a backreference to the first character group `\\1`. This will match a word like `anna` where the first letter and second letter are reversed to complete the second half of the word.

expression 3, `(..)\1`, accepts a character group with two digits `(..)` followed by a single backreference `\1` to that two-character string. This will match a pattern like `sasquatch\1` since when the string search finds the characters "ch" they are immediately repeated by the backreference. It will also match a pattern with a single character and two backreferences like `a\1\1`.

expression 4, `"(.).\\1.\\1"`, accepts a first character group `(.)` and then a single character `.` follwed by a backreference to the first character group `\\1`, followed by another single character `.`, followed by a final backreference to the first character group `\\1`. This will match a pattern such as `amana`.

expression 5, `"(.)(.)(.).*\\3\\2\\1"`, accepts three consecutive character groups `(.)(.)(.)` and applies three backreferences to the third, second and first character groups respectively `\\3\\2\\1`. This will match a word like 'redder' where the first three letters are reversed to complete the second half of the word.


#4 Construct regular expressions to match words that:

Start and end with the same character.

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
test2 <- c("kayak", "church", "discussion")

# first and last same
str_view(test2,"^(.).*\\1$", match = TRUE)

# repeated pair
str_view(test2,"(..).*\\1", match = TRUE)

# triple_threat
str_view(test2, "(.).*\\1.*\\1", match = TRUE)

```

